---
title: "chapter3.Rmd"
author: "Ella"
date: "7 helmikuuta 2017"
output: html_document
---



#3. Logistic Regression

###Task 2. Brief summary of the data

Here I will present the new data, that consists of background information and so forth, but the variables I will be focusing on mostly in this task is 'alc_use' and 'high_use'. But firstly I shall present all the variables:
```{r}
url <- read.table("http://s3.amazonaws.com/assets.datacamp.com/production/course_2218/datasets/alc.txt", sep=",", header=TRUE)
```



```{r}
str(url)
```

The data consists of variables as adress, that is defined as either urban (U) or rural (R), famsize measures how big the family is in 2 different groups (GT3, LT3), greater than 3 and less than 3. Medu and Fedu measures the mothers and fathers education, and Mjob and Fjob describes what kind of work the mother and father does. The codes G1, G2 and G3 describes the different grades in two school subjects (portugal and maths), in the different periods. 


Here I also want to present the dimensions of the data:

```{r}
dim(url)
```

As we can see, there are 35 different variables and 382 observations.


```{r}
library(tidyr); library(dplyr); library(ggplot2)
```

##Task 3. Numerical and graphical exploring variables


###Alcohol consumption variables
The two new variables I will be studying the relationship with other variables are alc_use and high_use. These variables consists of other variables combined. Alc_use consists of the avergage of the variables Dalc and Walc, that are the variables that measure how much alcohol is consumed during the workweek and the weekends. The high_use is when the alc_use on avreage is over 2. 

####Tables and visuals of the variable alc_use

```{r}
table(url$alc_use)
```

```{r}
g7 <- ggplot(data = url, aes(x = alc_use, fill = sex))
g7 + geom_bar()
```

####Tables and visuals of the variable high_use

```{r}
table(url$high_use)
```

```{r}
g8 <- ggplot(data = url, aes(x = high_use, fill = sex))
g8 + geom_bar()
```


###4 interesting variables
The first variable is 'failures', this variable measures how many times the individual has failed from a scale from 0-3 and more than 4. I think high use of alcohol could lead to a higher rate of failure in school. This bases on the hypothesis that alcohol consumption could lead to less success in school. 

  The second variable is 'absences' that measures how many times the person has been abscent from school. The variable ranges from 0-93.
  
  The third variable is 'famrel' that measures the quality of the relationship with the family. This variable ranges from 1-5, where 1 is very bad, and 5 is excellent.
  
  The forth variable 'goout' that measures on a scale from 1-5, very low(1) to very high (5). 



###Tables and visuals of the variable failures


```{r}
table(url$failures)
```
Just to visualize the failures variable, I will 
```{r}
g3 <- ggplot(data = url, aes(x = failures, fill = sex))
g3 + geom_bar()
```


###Tables and visuals of the variable absences


```{r}
table(url$absences)
```

```{r}
g4 <- ggplot(data = url, aes(x = absences, fill = sex))
g4 + geom_bar()
```


###Tables and visuals of the variable famrel

```{r}
table(url$famrel)
```

```{r}
g5 <- ggplot(data = url, aes(x = famrel, fill = sex))
g5 + geom_bar()
```

###Tables and visuals of the variable goout

```{r}
table(url$goout)
```

```{r}
g6 <- ggplot(data = url, aes(x = goout, fill = sex))
g6 + geom_bar()
```




###Exploring the relationships between the variables

####High_use vs absences

```{r}

g41 <- ggplot(url, aes(x = high_use, y = absences, col = sex))
g41 + geom_boxplot() + ylab("absences")

```

Here we can explore the individuals that has a high use of alcohol and the once whom not have a high use of alcohol.

We can see a slight difference between the women with high and low alcohol consumption in the difference in absences. But the biggest difference lies in the diffrence between absence with high and low use of alcohol between men. In the low use (FALSE) half of the men are distributed between 0 and less than 5 absences on average. On the other hand, in the high use(TRUE) we can see that half of this group has half of their distributions between 3 and 13 absences on average.  

