---
output:
  html_document: default
  pdf_document: default
---
# Insert chapter 2 title here

*Describe the work you have done this week and summarize your learning.*

- Describe your work and results clearly. 
- Assume the reader has an introductory course level understanding of writing and reading R code as well as statistical methods
- Assume the reader has no previous knowledge of your data or the more advanced methods you are using  
```{r}
data1 <- read.table("data/learning2014.txt" , sep = ",", 
header = TRUE)
```




I will begin to summaries the content of data1 and show its different dimensions:
```{r}
summary(data1)

```
```{r}
dim(data1)
```
As we can se, the data consists of 7 different variables gender, age, attitude, deep, stra, surf and points, and it has a total of 166 observations. The summary function gives 6 different basic key figures; minimum, 1st quartile, median, mean, 3rd quartile and maximum values.

Just to clarify how the data is structured:

```{r}
str(data1)
```

Just to visualize, I will have a picture of each variable, 

**-gender distribution**

```{r}
plot(data1$gender)
```
As we can se, a majority of the respondents are women, and from the summary we can find that there are 110 women and 56 men.


**-age distribution**

```{r}
plot(data1$age)
```
Here we can see that most of the respondents are in their 20s and the mean age is arounf 25.5 years. 


**-Attitude distribution**

```{r}
plot(data1$attitude)
```
Here we can se how the the attitude variable is very scattered, and has a median and mean around 3. This could indicate that people have very different attitudes regarding this surtain topic.

**-Deep variable**

```{r}
plot(data1$deep)
```
Here we can se that most of the respondents has a value between 3 and 4 (1st and 3d quartile) on the deep variable

**- Stra variable**

```{r}
plot(data1$stra)
```
The distribution in variable stra is also quite scattered, but the middle 50% is between 2,625 and 3,625 (1st and 3rd quartile)

**- Surf variable**

```{r}
plot(data1$surf)
```
In the surf variable we can see that the distribution is gathered between 2,4 and 3,4. 
**-points variable**

```{r}
plot(data1$points)
```
Here we can se that the points range between 7 and 33 points, and most of the respondents has over 20 points. 
